Estimation of the perceived time of presence of sources in urban acoustic environments using deep learning techniques

### 2019-11-26 paper/submitted/gontierAa2019v2.tex

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{pdflscape}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{bbm}
\usepackage{color}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\ml}[1]{\textcolor{red}{ML : #1}}
\begin{document}
\author{F\'elix Gontier $^1$, Catherine Lavandier $^2$, Pierre Aumond $^{2, 3}$\\Mathieu Lagrange $^1$, Jean-Francois Petiot $^1$}
\date{
$^1$ LS2N, UMR CNRS 6004, Ecole Centrale de Nantes, F-44321, France\\
$^2$ ETIS, UMR CNRS 8051, University of Paris Seine, University of Cergy-Pontoise, ENSEA, CNRS, F-95000, France\\
$^3$ IFSTTAR, CEREMA, UMRAE, F-44344, Bouguenais, France
}
\title{Estimation of the perceived time of presence of sources in urban acoustic environments using deep learning techniques}
\maketitle

@@

\begin{figure*}[th]
    \centering
     \begin{subfigure}[t]{0.33\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/tv_pres.eps}
    \end{subfigure}%
    \begin{subfigure}[t]{0.33\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/tb_pres.eps}
    \end{subfigure}
    \begin{subfigure}[t]{0.33\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/vb_pres.eps}
    \end{subfigure}
    \caption{Physical estimations~\cite{gontier2018} of the perceived time of presence of traffic, voices and birds sources for the corpus of simulated scenes with scenarios generated by \textit{simScene}.}\label{fig:tvb_pres}
\end{figure*}

@@

\begin{figure}[th]
    \centering
    \includegraphics[width=0.8\columnwidth]{figures/pca_p1.eps}
    \caption{Biplot of the principal components analysis of average assessments for the 5 general questions on the 6 recorded and 19 replicated scenes (n=25). Arrows show discrepancies between corresponding recorded and replicated scenes. For the P1 recording location the standard deviation of the projection of individual responses is shown using ellipses.}\label{fig:pspace_rec}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\columnwidth]{figures/pca_sim.eps}
    \caption{Biplot of the principal components analysis of average assessments for the 5 general questions on the 75 simulated scenes (n=75). Assessments of simulated scenes (active individuals) are projected as dots, and recorded and replicated scenes (supplementary individuals) are projected as crosses.}\label{fig:pspace_sim}
\end{figure}

@@

\begin{figure*}[t!]
    \centering
    \includegraphics[width=1.5\columnwidth]{figures/pred_ex.png}
    \caption{Example of simulated sound scene with ground truth source contributions and resulting presence predictions by the deep learning architecture (greyed out areas)}\label{fig:pred_ex}
\end{figure*}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{figures/deep_arch.pdf}
    \caption{Architecture of the deep learning model used for source presence prediction.}\label{fig:deep_arch}
\end{figure}

@@

\bibliographystyle{unsrt}
\bibliography{refs}

>>>>

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{pdflscape}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{bbm}
\usepackage{color}

\graphicspath{{../figures/}{./}}

\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\ml}[1]{\textcolor{red}{ML : #1}}
\begin{document}
\author{F\'elix Gontier $^1$, Catherine Lavandier $^2$, Pierre Aumond $^{2, 3}$\\Mathieu Lagrange $^1$, Jean-Francois Petiot $^1$}
\date{
$^1$ LS2N, UMR CNRS 6004, Ecole Centrale de Nantes, F-44321, France\\
$^2$ ETIS, UMR CNRS 8051, University of Paris Seine, University of Cergy-Pontoise, ENSEA, CNRS, F-95000, France\\
$^3$ IFSTTAR, CEREMA, UMRAE, F-44344, Bouguenais, France
}
\title{Estimation of the perceived time of presence of sources in urban acoustic environments using deep learning techniques}
\maketitle

@@

\begin{figure*}[th]
    \centering
     \begin{subfigure}[t]{0.33\textwidth}
        \centering
        \includegraphics[width=\textwidth]{tv_pres.eps}
    \end{subfigure}%
    \begin{subfigure}[t]{0.33\textwidth}
        \centering
        \includegraphics[width=\textwidth]{tb_pres.eps}
    \end{subfigure}
    \begin{subfigure}[t]{0.33\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vb_pres.eps}
    \end{subfigure}
    \caption{Physical estimations~\cite{gontier2018} of the perceived time of presence of traffic, voices and birds sources for the corpus of simulated scenes with scenarios generated by \textit{simScene}.}\label{fig:tvb_pres}
\end{figure*}

@@

\begin{figure}[th]
    \centering
    \includegraphics[width=0.8\columnwidth]{pca_p1.eps}
    \caption{Biplot of the principal components analysis of average assessments for the 5 general questions on the 6 recorded and 19 replicated scenes (n=25). Arrows show discrepancies between corresponding recorded and replicated scenes. For the P1 recording location the standard deviation of the projection of individual responses is shown using ellipses.}\label{fig:pspace_rec}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\columnwidth]{pca_sim.eps}
    \caption{Biplot of the principal components analysis of average assessments for the 5 general questions on the 75 simulated scenes (n=75). Assessments of simulated scenes (active individuals) are projected as dots, and recorded and replicated scenes (supplementary individuals) are projected as crosses.}\label{fig:pspace_sim}
\end{figure}

@@

\begin{figure*}[t!]
    \centering
    \includegraphics[width=1.5\columnwidth]{pred_ex.png}
    \caption{Example of simulated sound scene with ground truth source contributions and resulting presence predictions by the deep learning architecture (greyed out areas)}\label{fig:pred_ex}
\end{figure*}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{deep_arch.pdf}
    \caption{Architecture of the deep learning model used for source presence prediction.}\label{fig:deep_arch}
\end{figure}

@@

\begin{thebibliography}{10}

\bibitem{ardouin2018}
J.~Ardouin, L.~Charpentier, M.~Lagrange, F.~Gontier, N.~Fortin, J.~Picaut,
  D.~Ecoti\`ere, G.~Guillaume, and C.~Mietlicky.
\newblock An innovative low-cost sensor for urban sound monitoring.
\newblock In {\em INTER-NOISE 2018}, 2018.

\bibitem{mydlarz2017}
C.~Mydlarz, J.~Salamon, and J.P. Bello.
\newblock The implementation of low-cost urban acoustic monitoring devices.
\newblock {\em Applied Acoustics}, 117:207--218, 2017.

\bibitem{iso2014}
{ISO 12913-1:2014}.
\newblock Acoustics - soundscape - part 1: definition and conceptual framework.
\newblock Standard, International Organization for Standardization, Geneva, CH,
  2014.

\bibitem{can2008}
A.~Can, L.~Leclerq, J.~Lelong, and J.~Defrance.
\newblock Capturing urban traffic noise dynamics through relevant descriptors.
\newblock {\em Applied Acoustics}, 69:1270--1280, 2008.

\bibitem{brocolini2013}
L.~Brocolini, C.~Lavandier, M.~Quoy, and C.~Ribeiro.
\newblock Measurement of acoustic environments for urban soundscapes: choice of
  homogeneous periods, optimization of durations, and selection of indicators.
\newblock {\em J. Ac. Soc. Am.}, 134:813--821, 2013.

\bibitem{nilsson2006}
M.E. Nilsson and B.~Berglund.
\newblock Soundscape quality in suburban green areas and city parks.
\newblock {\em Acta Acust. unit. Acust.}, 92:903--911, 2006.

\bibitem{lunden2016}
P.~Lund\'en, O.~Axelsson, and M.~Hurtig.
\newblock On urban soundscape mapping: A computer can predict the outcome of
  soundscape assessments.
\newblock In {\em INTER-NOISE 2016}, 2016.

\bibitem{aucouturier2007}
J.-J. Aucouturier, B.~Defreville, and F.~Pachet.
\newblock The bag-of-frames approach to audio pattern recognition: a sufficient
  model for urban soundscapes but not for polyphonic music.
\newblock {\em J. Ac. Soc. Am.}, 122:881--91, 2007.

\bibitem{cakir2015}
E.~Cakir, T.~Heittola, H.~Huttunen, and T.~Virtanen.
\newblock Polyphonic sound event detection using multi label deep neural
  networks.
\newblock In {\em 2015 International Joint Conference on Neural Networks
  (IJCNN)}, 2015.

\bibitem{gontier2017}
F.~Gontier, M.~Lagrange, P.~Aumond, A.~Can, and C.~Lavandier.
\newblock An efficient audio coding scheme for quantitative and qualitative
  large scale acoustic monitoring using the sensor grid approach.
\newblock {\em Sensors}, 17, 2017.

\bibitem{aumond2017}
P.~Aumond, A.~Can, B.~De Coensel, D.~Botteldooren, C.~Ribeiro, and
  C.~Lavandier.
\newblock Modeling soundscape pleasantness using perceptive assessments and
  acoustic measurements along paths in urban context.
\newblock {\em Acta Acust. unit. Acust.}, 103:430--443, 2017.

\bibitem{torija2013}
A.~J. Torija, D.~P. Ruiz, and A.~F. Ramos-Ridao.
\newblock Application of a methodology for categorizing and differentiating
  urban soundscapes using acoustical descriptors and semantic-differential
  attributes.
\newblock {\em J. Ac. Soc. Am.}, 134:791--802, 2013.

\bibitem{nilsson2007}
M.E. Nilsson, D.~Botteldooren, and B.~De Coensel.
\newblock Acoustic indicators of soundscape quality and noise annoyance in
  outdoor urban areas.
\newblock In {\em 19th International Congress on Acoustics}, 2007.

\bibitem{viollon2000}
S.~Viollon and C.~Lavandier.
\newblock Multidimensional assessment of the acoustic quality of urban
  environments.
\newblock In {\em INTER-NOISE 2000}, 2000.

\bibitem{axelsson2010}
O.~Axelsson, M.E. Nilsson, and B.~Berglund.
\newblock A principal components model of soundscape perception.
\newblock {\em J. Ac. Soc. Am.}, 128:2836, 2010.

\bibitem{cain2013}
R.~Cain, P.~Jennings, and J.~Poxon.
\newblock The development and application of the emotional dimensions of a
  soundscape.
\newblock {\em Applied Acoustics}, 74:232--239, 2013.

\bibitem{jeon2018}
J.Y. Jeon, J.Y. Hong, C.~Lavandier, J.~Lafon, O.~Axelsson, and M.~Hurtig.
\newblock A cross-national comparison in assessment of urban park soundscapes
  in {F}rance, {K}orea, and {S}weden through laboratory experiments.
\newblock {\em Applied Acoustics}, 133:107--117, 2018.

\bibitem{aletta2016}
F.~Aletta, J.~Kang, and O.~Axelsson.
\newblock Soundscape descriptors and a conceptual framework for developing
  predictive soundscape models.
\newblock {\em Landsc. Urban Plan.}, 149:65--74, 2016.

\bibitem{delaitre2014}
P.~Delaitre, C.~Lavandier, C.~Ribeiro, M.~Quoy, E.~D'Hondt, E.~Gonzalez Boix,
  and K.~Kambona.
\newblock Influence of loudness of noise events on perceived sound quality in
  urban context.
\newblock In {\em Inter Noise}, 2014.

\bibitem{guastavino2007}
C.~Guastavino.
\newblock Categorization of environmental sounds.
\newblock {\em Can. J. Exp. Psychol.}, 61:54--63, 2007.

\bibitem{gygi2007}
B.~Gygi, G.~R. Kidd, and C.~S. Watson.
\newblock Similarity and categorization of environmental sounds.
\newblock {\em Perception \& Psychophysics}, 69:839--55, 2007.

\bibitem{brown2011}
A.L. Brown.
\newblock Towards standardization in soundscape preference assessment.
\newblock {\em Applied Acoustics}, 72:387--392, 2011.

\bibitem{lavandier2006}
C.~Lavandier and B.~Defreville.
\newblock The contribution of sound source characteristics in the assessment of
  urban soundscapes.
\newblock {\em Acta Acust. unit. Acust.}, 92:912--921, 2006.

\bibitem{ricciardi2014}
P.~Ricciardi, P.~Delaitre, C.~Lavandier, F.~Torchia, and P.~Aumond.
\newblock Sound quality indicators for urban places in paris cross-validated by
  {M}ilan data.
\newblock {\em J. Ac. Soc. Am.}, 138:2337--2348, 2014.

\bibitem{hong2017}
J.Y. Hong and J.Y. Jeon.
\newblock Relationship between spatiotemporal variability of soundscape and
  urban morphology in a multifunctional urban area: A case study in {S}eoul,
  {K}orea.
\newblock {\em Building and Environment}, 126:382--95, 2017.

\bibitem{lavandier2016}
C.~Lavandier, P.~Aumond, S.~Gomez, and C.~Domingu\`es.
\newblock Urban soundscape maps modelles with geo-referenced data.
\newblock {\em Noise Mapping}, 3:278--94, 2016.

\bibitem{mesaros2017}
A.~Mesaros, T.~Heitolla, A.~Diment, B.~Elizalde, A.~Shah, E.~Vincent, B.~Raj,
  and T.~Virtanen.
\newblock {DCASE} 2017 challenge setup: tasks, datasets and baseline system.
\newblock In {\em Detection and Classification of Acoustic Scenes and Events
  (DCASE) 2017 workshop}, 2017.

\bibitem{foggia2015}
P.~Foggia, N.~Petkov, A.~Saggese, N.~Strisciuglio, and M.~Vento.
\newblock Reliable detection of audio events in highly noisy environments.
\newblock {\em Pattern Recognition Letters}, 65:22--8, 2015.

\bibitem{mcfee2018}
B.~Mc{F}ee, J.~Salamon, and J.~P. Bello.
\newblock Adaptive pooling operators for weakly labeled sound event detection.
\newblock {\em IEEE Transactions on Audio, Speech and Language Processing},
  26:2180--93, 2018.

\bibitem{adavanne2017}
S.~Adavanne, P.~Pertila, and T.~Virtanen.
\newblock Sound event detection using spatial features and convolutional
  recurrent neural network.
\newblock In {\em 2017 IEEE International Conference on Audio, Speech and
  Signal Processing (ICASSP)}, 2017.

\bibitem{boes2018}
M.~Boes, K.~Filipan, B.~De Coensel, and D.~Botteldooren.
\newblock Machine listening for park soundscape quality assessment.
\newblock {\em Acta Acust. unit. Acust.}, 104:121--30, 2018.

\bibitem{barchiesi2015}
D.~Barchiesi, D.~Giannoulis, D.~Stowell, and M.~D. Plumbley.
\newblock Acoustic scene classification.
\newblock {\em IEEE Signal Processing Magazine}, 32:16--34, 2015.

\bibitem{sun2019}
K.~Sun, B.~De Coensel, K.~Filipan, F.~Aletta, T.~Van Renterghem, T.~De
  Pessemier, W.~Joseph, and D.~Botteldooren.
\newblock Classification of soundscapes of urban public open spaces.
\newblock {\em Landscape and Urban Planning}, 189:139--55, 2019.

\bibitem{valenti2016}
M.~Valenti, A.~Diment, G.~Parascandolo, S.~Squartini, and T.~Virtanen.
\newblock {DCASE} 2016 acoustic scene classification using convolutional neural
  networks.
\newblock In {\em Detection and Classification of Acoustic Scenes and Events
  (DCASE) 2016 workshop}, 2016.

\bibitem{salamon2017-2}
J.~Salamon and J.~P. Bello.
\newblock Deep convolutional neural networks and data augmentation for
  environmental sound classification.
\newblock {\em IEEE Signal Processing Letters}, 24, 2017.

\bibitem{yu2009}
L.~Yu and J.~Kang.
\newblock Modeling subjective evaluation of soundscape quality in urban open
  spaces: An artificial neural network approach.
\newblock {\em J. Ac. Soc. Am.}, 126:1163--74, 2009.

\bibitem{mesaros2019}
A.~Mesaros, A.~Diment, B.~Elizalde, T.~Heitolla, E.~Vincent, B.~Raj, and
  T.~Virtanen.
\newblock Sound event detection in the {DCASE} 2017 challenge.
\newblock {\em IEEE Transactions on Audio, Speech and Language Processing},
  27:992--1006, 2019.

\bibitem{salamon2014}
J.~Salamon, C.~Jacoby, and J.~P. Bello.
\newblock A dataset and taxonomy for urban sound research.
\newblock In {\em 22nd ACM International Conference on Multimedia}, 2014.

\bibitem{gemmeke2017}
J.~F. Gemmeke, D.~Ellis, D.~Freedman, A.~Jansen, W.~Lawrence, R.~Moore,
  M.~Plakal, and M.~Ritter.
\newblock Audio {S}et: an ontology and human-labeled dataset for audio events.
\newblock In {\em 2017 IEEE International Conference on Audio, Speech and
  Signal Processing (ICASSP)}, 2017.

\bibitem{lafay2016}
Gr{\'e}goire Lafay, Mathieu Lagrange, Mathias Rossignol, Emmanouil Benetos, and
  Axel Roebel.
\newblock {A morphological model for simulating acoustic scenes and its
  application to sound event detection}.
\newblock {\em {IEEE/ACM Transactions on Audio, Speech and Language
  Processing}}, 24(10):1854--1864, 2016.

\bibitem{salamon2017}
J.~Salamon, D.~MacConnell, M.~Cartwright, P.~Li, and J.~P. Bello.
\newblock Scaper: A library for soundscape synthesis and augmentation.
\newblock In {\em 2017 IEEE Workshop on Applications of Signal Processing to
  Audio and Acoustics (WASPAA)}, 2017.

\bibitem{panayotov2015}
V.~Panayotov, G.~Chen, D.~Povey, and S.~Khudanpur.
\newblock Librispeech: an {ASR} corpus based on public domain audio books.
\newblock In {\em 2015 IEEE International Conference on Acoustics, Speech and
  Signal Processing (ICASSP)}, 2015.

\bibitem{gloaguen2017}
J.R. Gloaguen, A.~Can, M.~Lagrange, and J.F. Petiot.
\newblock Creation of a corpus of realistic urban sound scenes with controlled
  acoustic properties.
\newblock In {\em Meetings on Acoustics}, 2017.

\bibitem{gontier2018}
F.~Gontier, P.~Aumond, M.~Lagrange, C.~Lavandier, and J.-F. Petiot.
\newblock Towards perceptual soundscape characterization using event detection
  algorithms.
\newblock In {\em Detection and Classification of Acoustic Scenes and Events
  (DCASE) 2018 workshop}, 2018.

\bibitem{dagnelie2003}
P.~Dagnelie.
\newblock {\em Principes d'exp\'erimentation: planification des exp\'eriences
  et analyse de leurs r\'esultats}.
\newblock Presses Agronomiques de Gembloux, 2003.

\bibitem{moller1992}
H.~M{\o}ller.
\newblock Fundamentals of binaural technology.
\newblock {\em Applied Acoustics}, 36:171--218, 1992.

\bibitem{wilcoxon1945}
F.~Wilcoxon.
\newblock Individual comparisons by ranking methods.
\newblock {\em Biometrics Bulletin}, 1:80--3, 1945.

\bibitem{pratt1959}
J.~W. Pratt.
\newblock Remarks on zeros and ties in the {W}ilcoxon signed rank procedures.
\newblock {\em Journal of the American Statistical Association}, 54:655--67,
  1959.

\bibitem{itatoolbox2017}
M.~Berzborn, R.~Bomhardt, J.~Klein, J.G. Richter, and M.~Vorlander.
\newblock The ita-toolbox: an open source matlab toolbox for acoustic
  measurements and signal processing.
\newblock In {\em 43th Annual German Congress on Acoustics}, 2017.

\bibitem{paszke2017}
A.~Paszke, S.~Gross, S.~Chintala, G.~Chanan, E.~Yang, Z.~DeVito, Z.~Lin,
  A.~Desmaison, L.~Antiga, and A.~Lerer.
\newblock Automatic differentiation in pytorch.
\newblock In {\em 2017 Conference Neural Information Processing Systems
  (NIPS)}, 2017.

\bibitem{kingma2015}
D.~P. Kingma and J.~Lei Ba.
\newblock Adam: a method for stochastic optimization.
\newblock In {\em 2015 International Conference on Learning Representations
  (ICLR)}, 2015.

\bibitem{blauert1997}
J.~Blauert and U.~Jekosch.
\newblock Sound quality evaluation: a multi layered problem.
\newblock {\em Acta Acust. unit. Acust.}, 83:747--53, 1997.

\bibitem{jekosch2004}
U.~Jekosch.
\newblock Basic concepts and terms quality, reconsidered in the context of
  product-sound quality.
\newblock {\em Acta Acust. unit. Acust.}, 90:999--1006, 2004.

\bibitem{lasseck2018}
M.~Lasseck.
\newblock Acoustic bird detection with deep convolutional neural networks.
\newblock In {\em Detection and Classification of Acoustic Scenes and Events
  (DCASE) 2016 workshop}, 2018.

\end{thebibliography}
